---
title: "Microbial Carbon Traits"
author: "Mario E. Muscarella"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
   - \usepackage{graphics}
output: 
  pdf_document:
    fig_caption: true
---

# Initial Setup
```{r results='hide', message=FALSE}
rm(list=ls())
getwd()
setwd("~/GitHub/MicrobialCarbonTraits/analyses")
se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
sem <- function(x){sd(na.omit(x))/sqrt(length(na.omit(x)))}

# Save Default Plot Settings
opar <- par(no.readonly = TRUE)  # Saves plot defaults

# Required Packages
#require("muscle");require("seqinr");require("caper")
require("ape");require("phylobase");require("adephylo")
require("geiger");require("picante")
require("stats");require("RColorBrewer")
require("png"); require("grid");require("reshape"); require("car")
```

# Isolate Information
## Figure 1: Phylogenetic Tree
Do I want to make a really nice tree or just use the species tree in the next figure
```{r, results = "hide"}
# Import Phylogenetic Tree



```

# Bacterial Growth Efficiency
```{r}
# Import Data
BP <- read.csv(file="../data/CarbonTraits/BP_data.txt")
BR <- read.csv(file="../data/CarbonTraits/BR_data.txt")

BP$Organism == BR$Organism

BGE <- as.data.frame(matrix(NA, 23, 4))
colnames(BGE) <- c("Organism", "BGE_Glu", "BGE_Suc", "BGE_Pro")
BGE$Organism <- BP$Organism

BGE$BGE_Glu <- round(BP$Glu_mean / (BP$Glu_mean + BR$Glu_mean), 5)
BGE$BGE_Suc <- round(BP$Suc_mean / (BP$Suc_mean + BR$Suc_mean), 5)
BGE$BGE_Pro <- round(BP$Pro_mean / (BP$Pro_mean + BR$Pro_mean), 5)

# Import Phylogeny (already rooted)
tree <- read.tree("../data/Phylogeny/HMWF.proteo.nwk.tre")

# Keep Rooted but Drop Outgroup Branch
tree <- drop.tip(tree, "Aquifex")
tree$edge.length <- tree$edge.length + 10^-7

# Carbon Respiration Traits {adephylo}
Resp <- BR[,c(11, 13, 15)]
Resp[is.na(Resp)] <- 0
Resp[Resp > 30] <- 30
colnames(Resp) <- c("Glucose", "Succinate", "Protocatechuate")
rownames(Resp) <- BR$Organism

# Reorder Resp
Resp <- as.matrix(Resp[match(tree$tip.label, row.names(Resp)), ])

# Carbon Production Traits
Prod <- BP[,c(11, 13, 15)]
Prod[is.na(Prod)] <- 0
Prod[Prod < 0] <- 0
Prod[Prod > 10] <- 10
colnames(Prod) <- c("Glucose", "Succinate", "Protocatechuate")
rownames(Prod) <- BP$Organism

# Reorder Prod
Prod <- as.matrix(Prod[match(tree$tip.label, row.names(Prod)), ])

# Carbon BGE Traits
BGE2 <- BGE[,2:4]
BGE2[is.na(BGE2)] <- 0
BGE2[BGE2 < 0] <- 0
BGE2[BGE2 > 0.9] <- 0
colnames(BGE2) <- c("Glu.", "Suc.", "Pro.")
rownames(BGE2) <- BGE$Organism

# Reorder BGE
BGE2 <- as.matrix(BGE2[match(tree$tip.label, row.names(BGE2)), ])
```

##Figure 2: Bacterial Growth Efficiency
```{r, results='hide'}
png(filename="../figures/Figure2.png",
    width = 1200, height = 1200, res = 96*2, bg = "white")
par(opar)

# Define Color Palette
mypalette <- colorRampPalette(brewer.pal(9, "YlOrRd"), bias = 2)

# Make Plot
par(mar=c(1,1,1,7) + 0.1)
x <- phylo4d(tree, tip.data = BGE2)
table.phylo4d(x, treetype = "phylo", symbol = "colors", show.node = FALSE,
              cex.label = 0.8, scale = FALSE, use.edge.length = FALSE,
              edge.color = "black", edge.width = 3, box = FALSE,
              col=mypalette(25), pch = 15, cex.symbol = 2.5,
              ratio.tree = 0.7, cex.legend = 1.25, center = FALSE,
              show.tip.label = FALSE)
mtext(tipLabels(x), side =4, las = 1, at = c(1:23), cex = 0.8)
text(18.5, 22.5, expression(alpha), cex=1.5)
text(18.5, 19.5, expression(beta), cex=1.5)
text(4, 13, expression(gamma), cex=1.5)
text(11.5, 17, "Xan.", cex=1)
text(11.5, 12, "Aero.", cex=1)
text(11.5, 7.5, "Pseu.", cex=1)


dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
```

```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="BGE"}
img <- readPNG("../figures/Figure2.png")
grid.raster(img)
```

## Statistical Tests
```{r}
# Hypothesis Testing
traits <- cbind(Resp, Prod, BGE2)

# Import Phylogeny (already rooted)
tree1000 <- read.tree("../data/Phylogeny/RAxML_1000bootstrap.HMWF.tre")

# Reorder Traits
tree1000[[1]] <- drop.tip(tree1000[[1]], "Aquifex")
traits <- as.matrix(traits[match(tree1000[[1]]$tip.label, row.names(traits)), ])

# Blomberg's K - BGE
# bloms3 <- matrix(NA, 1000, 6)
# colnames(bloms3) <- c("K_g", "K_s", "K_p", "P_g", "P_s", "P_p")
# for (i in 1:3){
#   for (j in 1:1000){
#     tree1000[[j]] <- drop.tip(tree1000[[j]], "Aquifex")
#     tree1000[[j]]$edge.length <- tree1000[[j]]$edge.length + 10^-7
#     output <- phylosignal(traits[, 6 + i], tree1000[[j]])
#     bloms3[j, i] <- output[[1]]
#     bloms3[j, 3+i] <- output[[4]]
#   }}
# 
# avgK1 <- colMeans(bloms3)[1:3]
# avgP <- colMeans(bloms3)[4:6]
# length(which(bloms3[, 4] > 0.05)) / 1000
# length(which(bloms3[, 5] > 0.05)) / 1000
# length(which(bloms3[, 5] > 0.05)) / 1000
# 
# 
# phylosignal(traits[, 6 + i], tree)
# 

# Varaince Partitioning
BGE3 <- as.data.frame(BGE2)
BGE3$ID <- row.names(BGE3)
BGE3$Taxon <- c("Pseudo", "Pseudo", "Pseudo", "Pseudo", "Pseudo", "Pseudo", "Pseudo",
                "Pseudo", "Pseudo", "Aero", "Aero", "Aero", "Aero", "Aero", "Xanto",
                "Xanto", "Xanto", "Beta", "Beta", "Beta", "Alpha", "Alpha", "Alpha")
BGE3$TaxonP <- c("Gamma", "Gamma", "Gamma", "Gamma", "Gamma", "Gamma", "Gamma",
                "Gamma", "Gamma", "Gamma", "Gamma", "Gamma", "Gamma", "Gamma", "Gamma",
                "Gamma", "Gamma", "Beta", "Beta", "Beta", "Alpha", "Alpha", "Alpha")



# Mixed Model
BGE.l <- melt(BGE3)
model <- glm(log1p(BGE.l$value) ~ as.factor(BGE.l$Taxon) + as.factor(BGE.l$variable))
summary(model)

1 - (0.29902/0.40510)
1 - (0.35154/0.40510)

Anova(model, type="III")


# Mixed Model
BGE.l <- melt(BGE3)
model <- glm(log1p(BGE.l$value) ~ as.factor(BGE.l$Taxon))
summary(model)

1 - (0.32573/0.40510)

# Mixed Model
BGE.l <- melt(BGE3)
model <- glm(log1p(BGE.l$value) ~ as.factor(BGE.l$TaxonP))
summary(model)

1 - (0.37824/0.40510)

Anova(model, type="III")


# Mixed Model
BGE.l <- melt(BGE3)
model <- glm(log1p(BGE.l$value) ~ as.factor(BGE.l$variable))
summary(model)
1 - (0.3784/0.40510)

Anova(model, type="III")





# Generate Test Statistics for Comparing Phylogenetic Signal {geiger}
#fitContinuous(nj.rooted, nb, model = "lambda")
#fitContinuous(nj.lambda.0, nb, model = "lambda")



# First, Correct for Zero Branch-Lengths on Our Tree
#nj.rooted$edge.length <- nj.rooted$edge.length + 10^-7

# Calculate Phylogenetic Signal for Growth on All Phosphorus Resources
# First, Create a Blank Output Matrix
#p.phylosignal <- matrix(NA, 6, 18)
#colnames(p.phylosignal) <- colnames(p.growth.std)
#rownames(p.phylosignal) <- c("K", "PIC.var.obs", "PIC.var.mean",
#                             "PIC.var.P", "PIC.var.z", "PIC.P.BH")

# Use a For Loop to Calculate Blomberg's K for Each Resource
#for (i in 1:18){
#  x <- as.matrix(p.growth.std[ ,i, drop = FALSE])
#  out <- phylosignal(x, nj.rooted)
#  p.phylosignal[1:5, i] <- round(t(out), 3)
#}

# Use the BH Correction on P-values:
#p.phylosignal[6, ] <- round(p.adjust(p.phylosignal[4, ], method = "BH"), 3)


# Calcualate Phylogenetic Signal for Niche Breadth
#signal.nb <- phylosignal(nb, nj.rooted)



```

# BGE under the Generalist-Specialist Framework
```{r}
# Import Data
ecoplate <- read.delim("../data/EcoPlate/eco.data.txt")
pathways <- read.delim("../data/Maple/pathways.txt")

levins <- function(p_xi = ""){
  p = 0
  for (i in p_xi){
    p = p + i^2
    }
  nb = 1 / (length(p_xi) * p)
  return(nb)
}

ecoplate.std <- ecoplate[,2:32] / (apply(ecoplate[,2:32], 1, sum))

ecoplate$levins <- levins(ecoplate.std)

ecoplate2 <- ecoplate[ecoplate$Strain %in% BGE$Organism, ]
pathways2 <- pathways[pathways$Organism %in% BGE$Organism, ]

niche <- merge(pathways2, ecoplate2[,c(1,33,34)], 
               by.x = "Organism", by.y = "Strain", all.x = T)

niche.bge <- merge(BGE, niche)
```





# BGE under the Copiotroph-Oligotroph Framework
```{r}
umax <- read.csv("../data/GrowthCurves/umax.txt", header=T)
copies <- read.delim("../data/CopyNumber/16SrRNA.txt")

growth <- merge(umax, copies, by.x = "isolate", by.y = "Organism")

growth.bge <- merge(BGE, growth, by.x = "Organism", by.y = "isolate")



```


# Figure 3: Species Traits
```{r, results='hide'}
png(filename="../figures/Figure3.png",
    width = 1200, height = 1200, res = 96*2, bg = "white")
par(opar)

layout(matrix(1:4, nrow = 2))

par(mar = c(2,1,1,1) + 0.5, oma = c(1, 3, 2, 1))

plot(BGE_Glu ~ levins, niche.bge, axes = "n", 
     xlim = c(0, 0.5), ylim = c(0, 0.35), las = 1,
     pch = 22, bg = "gray", lwd = 2, cex = 1.5)
axis(1, lwd = 2, labels = T, at = c(seq(0, 0.5, 0.1)), las = 1)
axis(2, lwd = 2, labels = T, at = c(0, 0.1, 0.2, 0.3), las = 1)
axis(1, lwd = 2, labels = F, tck = 0.02, at = c(seq(0, 0.5, 0.1)), las = 1)
axis(2, lwd = 2, labels = F, tck = 0.02, at = c(0, 0.1, 0.2, 0.3), las = 1)
axis(3, lwd = 2, labels = F, tck = -0.02, at = c(seq(0, 0.5, 0.1)), las = 1)
axis(4, lwd = 2, labels = F, tck = -0.02, at = c(0, 0.1, 0.2, 0.3), las = 1)
axis(3, lwd = 2, labels = F, tck = 0.02, at = c(seq(0, 0.5, 0.1)), las = 1)
axis(4, lwd = 2, labels = F, tck = 0.02, at = c(0, 0.1, 0.2, 0.3), las = 1)
mtext("Levins", side = 1, line = 2)
mtext("Generalist-Specialist", side = 3, line = 1)
box(lwd = 2)


plot(BGE_Glu ~ Pathways, niche.bge, axes = "n", 
     xlim = c(35, 75), ylim = c(0,0.35), las = 1,
     pch = 22, bg = "gray", lwd = 2, cex = 1.5)
axis(1, lwd = 2, labels = T, at = c(seq(40, 70, 10)), las = 1)
axis(2, lwd = 2, labels = T, at = c(0, 0.1, 0.2, 0.3), las = 1)
axis(1, lwd = 2, labels = F, tck = 0.02, at = c(seq(40, 70, 10)), las = 1)
axis(2, lwd = 2, labels = F, tck = 0.02, at = c(0, 0.1, 0.2, 0.3), las = 1)
axis(3, lwd = 2, labels = F, tck = -0.02, at = c(seq(40, 70, 10)), las = 1)
axis(4, lwd = 2, labels = F, tck = -0.02, at = c(0, 0.1, 0.2, 0.3), las = 1)
axis(3, lwd = 2, labels = F, tck = 0.02, at = c(seq(40, 70, 10)), las = 1)
axis(4, lwd = 2, labels = F, tck = 0.02, at = c(0, 0.1, 0.2, 0.3), las = 1)
mtext("Pathways", side = 1, line = 2)
box(lwd = 2)

plot(BGE_Glu ~ umax, growth.bge, axes = "n", 
     xlim = c(0, 0.2), ylim = c(0,0.35), las = 1,
     pch = 22, bg = "gray", lwd = 2, cex = 1.5)
axis(1, lwd = 2, labels = T, at = c(seq(0, 0.2, 0.05)), las = 1)
axis(2, lwd = 2, labels = T, at = c(0, 0.1, 0.2, 0.3), las = 1)
axis(1, lwd = 2, labels = F, tck = 0.02, at = c(seq(0, 0.2, 0.05)), las = 1)
axis(2, lwd = 2, labels = F, tck = 0.02, at = c(0, 0.1, 0.2, 0.3), las = 1)
axis(3, lwd = 2, labels = F, tck = -0.02, at = c(seq(0, 0.2, 0.05)), las = 1)
axis(4, lwd = 2, labels = F, tck = -0.02, at = c(0, 0.1, 0.2, 0.3), las = 1)
axis(3, lwd = 2, labels = F, tck = 0.02, at = c(seq(0, 0.2, 0.05)), las = 1)
axis(4, lwd = 2, labels = F, tck = 0.02, at = c(0, 0.1, 0.2, 0.3), las = 1)
mtext("Growth Rate", side = 1, line = 2)
mtext("Oligotroph-Copiotroph", side = 3, line = 1)
box(lwd = 2)


plot(BGE_Glu ~ CopyNumber, growth.bge, axes = "n",
     xlim = c(1, 11), ylim = c(0, 0.35),
     pch = 22, bg = "gray", lwd = 2, cex = 1.5)
axis(1, lwd = 2, labels = T, at = c(seq(2, 10, 2)), las = 1)
axis(2, lwd = 2, labels = T, at = c(0, 0.1, 0.2, 0.3), las = 1)
axis(1, lwd = 2, labels = F, tck = 0.02, at = c(seq(2, 10, 2)), las = 1)
axis(2, lwd = 2, labels = F, tck = 0.02, at = c(0, 0.1, 0.2, 0.3), las = 1)
axis(3, lwd = 2, labels = F, tck = -0.02, at = c(seq(2, 10, 2)), las = 1)
axis(4, lwd = 2, labels = F, tck = -0.02, at = c(0, 0.1, 0.2, 0.3), las = 1)
axis(3, lwd = 2, labels = F, tck = 0.02, at = c(seq(2, 10, 2)), las = 1)
axis(4, lwd = 2, labels = F, tck = 0.02, at = c(0, 0.1, 0.2, 0.3), las = 1)
mtext("Copy Number", side = 1, line = 2)
box(lwd = 2)

mtext("Bacterial Growth Efficiency", side = 2, line = 1, outer = T, cex = 1.5)


dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
#```

#```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Traits"}
img <- readPNG("../figures/Figure3.png")
grid.raster(img)
```



# BGE as a Genomic Constraints
```{r}
maple <- read.delim("../data/Maple/maple.txt", header = T, row.names = 1)
maple <- (maple > 80) * 1
maple <- maple[BGE$Organism,which(colSums(maple) > 1)]

maple2 <- maple[-c(which(is.na(BGE$BGE_Glu))),]
maple2 <- maple2[,which(colSums(maple2) > 1)]

# Calculate Bray-Curtis Distances for Genomes
maple.bray <- vegdist(maple, "bray")

# BGE dbRDA 
BGE.dbRDA <- capscale(maple.bray ~ BGE$BGE_Glu, comm = maple, 
                      add = T, na.action = na.exclude)
str(BGE.dbRDA$CCA$wa)
str(maple2)


# DOM Scores
pathway.scores <- data.frame(matrix(NA, 86, 2))
colnames(pathway.scores) <- c("pathway", "cor")
pathway.scores[,1] <- colnames(maple2)
for (i in 1:86){
  pathway.scores[i,2] <- cor(BGE.dbRDA$CCA$wa,maple2[,i], method = "spearman")
}

sig.path <- pathway.scores[which(abs(pathway.scores$cor) > 0.70), ]

plot(BGE.dbRDA$CCA$wa, na.omit(BGE$BGE_Glu))

# M00169	Pathway	CAM (Crassulacean acid metabolism), light
# M00045	Pathway	Histidine degradation, histidine => N-formiminoglutamate => glutamate
# M00124	Pathway	Pyridoxal biosynthesis, erythrose-4P => pyridoxal-5P
# M00133	Pathway	Polyamine biosynthesis, arginine => agmatine => putrescine => spermidine

sum(maple2[,"M00169"] > 0)
plot(maple2[,"M00169"],na.omit(BGE$BGE_Glu))

sum(na.omit(BGE$BGE_Glu) < 0.01)

```





